---
title: "Group Project: Project plan & Initial Data Analysis (IDA)"
subtitle: "STAT5003"
author: "Group 38"
params:
  soln: FALSE   ## EDIT to TRUE when generating solution
  supp: FALSE
  show: 'hide'  ## EDIT to 'as.is' when generating solutions, otherwise 'hide'
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
  pdf_document:
    number_sections: yes
    toc: yes
---

```{r setup, include=FALSE}
# For solutions
chooseCRANmirror(graphics=FALSE, ind=1)
knitr::opts_chunk$set(echo = TRUE,tidy=TRUE)
show_q = FALSE
show_s = FALSE
# Then include solutions file
# ```{r child='Lab1_Quiz_s.Rmd', eval = show_s}
```

Members:
<ul>
  <li>Quynh Giang Le - 530190754</li>
  <li>Kairui Huang - 530261917</li>
  <li>Rundong Shen - 530023418</li>
  <li>Haoyu Fang - 540576287</li>
  <li>Wanxin Bao - 540320288</li>
</ul>

```{r}
suppressPackageStartupMessages(library('tidyverse'))
suppressPackageStartupMessages(library('readr'))
suppressPackageStartupMessages(library('dplyr'))
suppressPackageStartupMessages(library('ggplot2'))
suppressPackageStartupMessages(library('broom'))
suppressPackageStartupMessages(library('tinytex'))
suppressPackageStartupMessages(library('caret'))
suppressPackageStartupMessages(library('gridExtra'))

data <- as.data.frame(read.csv("credit_train.csv", header = TRUE))
```

This dataset is from 
<br>Rohan Paris. (2022). Credit score classification [dataset]. Kaggle. https://www.kaggle.com/datasets/parisrohan/credit-score-classification

# Problem Overview {.tabset .tabset-fade .tabset-pills}
The purpose of this task is to construct a credit rating classification model based on existing data from banks, such as loan quantity, number of delayed payments, etc. After receiving user personal data, users are divided into pre-set different credit risk levels. This is a supervised learning task, where the training data has a known label, namely Credit_Score, and the label is discrete, making it a classification problem.<br>
<br>This task is interesting and has practical significance. Credit scoring is widely used in banks and other financial institutions. By constructing a classification model for credit scoring, it can effectively improve classification efficiency and reduce labor costs. When dealing with large amounts of lending data, this model will be particularly effective. At the same time, the model also improves the standardization of classification, ensuring the accuracy and objectivity of credit ratings, which helps credit ratings be accepted by consumers. From a technical perspective, building an efficient and accurate scoring model poses many challenges, including data cleaning, feature selection, model optimization, and so on. This task can help beginners master the basic knowledge of data science, understand the advantages and disadvantages of different models, and explore how to improve the performance of models. This is both a challenge and an interesting research process.

# Dataset Description {.tabset .tabset-fade .tabset-pills}

## Description
The dataset contains 100000 samples and 28 feature columns. Among these features, 16 columns represent numerical data and 12 columns represent categorical data. Among them, it should be noted that due to data entry or issues with outliers, some numerical information is saved in char format and needs to be cleaned and formatted, such as in the Num of Loan column. The column names and data types are shown in the following figure:

```{r}
str(data)
```
Among these variables, there are 17 numerical variables, including "Age", "Annual_Income", "Monthly_Inhand_Salary", "Num_Bank_Accounts", "Num_Credit_Card", "Interest_Rate", "Num_of_Loan", "Delay_from_due_date", "Num_of_Delayed_Payment", "Changed_Credit_Limit", "Num_Credit_Inquiries", "Outstanding_Debt", "Credit_Utilization_Ratio", "Credit_History_Age", "Total_EMI_per_month", "Amount_invested_monthly", and "Monthly_Balance", and 7 categorical variables, including "Month", "Occupation", "Credit_Mix", "Payment_of_Min_Amount", "Payment_Behaviour", and "Credit_Score". These variables are intended to be used to inspect the classfication model. For the other columns, they are unnecessary in performing the model or data cleaning.

## Dataset challenges
In this task, the dataset contains 100000 samples, with a large number of missing values that need to be supplemented in different ways. Filling in a large number of missing values will be a challenge. At the same time, it is necessary to handle various types of erroneous values during research, such as underline, values that clearly exceed the normal range, etc. This will involve a series of tasks such as data cleaning and data type conversion. Meanwhile, the model needs to solve a Multi class classification problem, which is also a challenge.

# Data cleaning {.tabset .tabset-fade .tabset-pills}

## Initial cleaning

<strong>1. Check for duplicated rows</strong>

```{r}
sum(duplicated(data))
```

<strong>2. Remove unnecessary columns</strong>

```{r}
data <- data[ , -which(names(data) %in% c("ID", "Name", "SSN"))]
```

<strong>3. Change value for Credit_History_Age column</strong>

```{r}
data["Credit_History_Age"] = lapply(data["Credit_History_Age"], function(x) {
    years <- as.integer(sub("\\Years.*", "", x))
    months <- as.integer(sub(".*\\d+\\s+Years\\s+and\\s+(\\d+)\\s+Months.*", "\\1", x))
    return(years*12 + months)
})
```

## Numerical variables

In this section, we would perform some visualization to see the distribution of some numerical variables, some checking functions to comprehend the pattern of these variables. As the result, we would do some modification to missing values and inappropriate outliers.

<strong>1. Delete leading/trailing "_" in some columns</strong>

```{r}
for(i in c("Age","Annual_Income", "Num_of_Loan", "Num_of_Delayed_Payment", 
                    "Changed_Credit_Limit", "Outstanding_Debt", "Amount_invested_monthly", "Monthly_Balance")) {
  data[[i]] <- gsub("_", "", data[[i]])
  
  data[[i]] <- as.numeric(data[[i]])
}
```

<strong>2. Declare supporting functions</strong>

```{r}
# This is a function to find the bound of the column by using IQR
calculate_bounds <- function(column_name) {
  Q1 <- quantile(data[[column_name]], 0.25, na.rm = TRUE)
  Q3 <- quantile(data[[column_name]], 0.75, na.rm = TRUE)
  IQR <- Q3 - Q1
  lower_bound <- Q1 - 1.5 * IQR
  upper_bound <- Q3 + 1.5 * IQR
  return(c(lower_bound, upper_bound))
}

# This is a function to check if each group has a unique value
analyze_unique_values <- function(income_column) {
  lower_bound <- calculate_bounds(income_column)[1]
  upper_bound <- calculate_bounds(income_column)[2]
  
  filtered_data <- data %>%
    filter(!!sym(income_column) >= lower_bound & !!sym(income_column) <= upper_bound)
  
  result <- filtered_data %>%
    group_by(Customer_ID) %>%
    summarise(Unique_Value_Count = n_distinct(!!sym(income_column)),
              Only_One_Unique_Value = Unique_Value_Count == 1) %>%
    ungroup()

  if(all(result$Only_One_Unique_Value)) {
    return (TRUE)
  } else {
    return (FALSE)
  }
}

# This is a function to change outliers with neighbor
clean_data_with_neighbor <- function(column_name, lower_bound, upper_bound) {
  column_sym <- sym(column_name)
  result <- data %>%
    group_by(Customer_ID) %>%
    mutate(
      !!column_sym := ifelse(
        .data[[column_name]] < lower_bound | .data[[column_name]] > upper_bound, NA, .data[[column_name]]
      )
    ) %>%
    fill(!!column_sym, .direction = "up") %>%
    fill(!!column_sym, .direction = "down") %>%
    ungroup()
  
  return(result)
}

# This is a function that can find the most common value in every column
find_mode <- function(x) {
  ux <- unique(na.omit(x))
  ux[which.max(tabulate(match(x, ux)))]
}

# This is a function to change outliers with mode
clean_data_with_mode <- function(column_name) {
  col_sym <- rlang::sym(column_name) 
  result <- data %>%
    group_by(Customer_ID) %>%
    mutate(Mode = find_mode(!!col_sym),
           !!col_sym := Mode) %>% 
    select(-Mode) %>%
    ungroup()
  return(result)
}

# This is a function to fill NA with neighbor
fill_NA_with_neighbour <- function(column_name) {
  result <- data %>%
    group_by(Customer_ID) %>%
    fill(!!rlang::sym(column_name), .direction = "up") %>%
    fill(!!rlang::sym(column_name), .direction = "down") %>%
    ungroup()
  
  return(result)
}

# This is a function to replace negative value with 0
replace_negative_with_0 <- function(column_name) {
  result <- data %>%
    mutate(!!rlang::sym(column_name) := ifelse(!!rlang::sym(column_name) < 0, 0, !!rlang::sym(column_name)))
  
  return(result)
}

# This is a function to fill NA with mean
fill_NA_with_mean <- function(column_name) {
  col_sym <- rlang::sym(column_name)
  result <- data %>%
    group_by(Customer_ID) %>%
    mutate(
      Mean = round(mean({{col_sym}}[!is.na({{col_sym}})], na.rm = TRUE), 2),
      {{col_sym}} := if_else(is.na({{col_sym}}), Mean, {{col_sym}})
    ) %>%
    select(-Mean) %>%
    ungroup()
  return(result)
}

# This is a function to replace the outliers and NA with mean
replace_with_group_mean <- function(column_name, value_to_replace) {
  col_sym <- sym(column_name)
  result <- data %>%
    group_by(Customer_ID) %>%
    mutate(
      Group_Mean = mean({{col_sym}}[{{col_sym}} != value_to_replace & !is.na({{col_sym}})], na.rm = TRUE),
      {{col_sym}} := if_else({{col_sym}} == value_to_replace | is.na({{col_sym}}), Group_Mean, {{col_sym}})
    ) %>%
    select(-Group_Mean) %>%
    ungroup()
  return(result)
}
```

<strong>3. Check the values of all numerical variables</strong>

<strong>3.1. Check missing value percentages</strong>

```{r}
missing_values <- colSums(is.na(data[c("Age", "Annual_Income", "Monthly_Inhand_Salary", "Num_Bank_Accounts", 
                                       "Num_Credit_Card", "Interest_Rate", "Num_of_Loan", 
                                       "Delay_from_due_date", "Num_of_Delayed_Payment", 
                                       "Changed_Credit_Limit", "Num_Credit_Inquiries", 
                                       "Outstanding_Debt", "Credit_Utilization_Ratio", 
                                       "Credit_History_Age", "Total_EMI_per_month", 
                                       "Amount_invested_monthly", "Monthly_Balance")]))


total_values <- nrow(data)

missing_values_percent <- missing_values / total_values * 100

missing_values_percent
```
<br> <strong> <u>Interpretation:</u></strong>
<br> Based on these results, there are not many columns have NA values; however, we would still do some further investigation on distribution to see if there is any invalid values or abnormal outliers

<br><br><strong>3.2. Histogram distribution of some columns</strong>

```{r}
hist_num_cols <- c("Age", "Num_Bank_Accounts", "Num_Credit_Card", "Num_of_Delayed_Payment", "Num_Credit_Inquiries", "Total_EMI_per_month", "Amount_invested_monthly", "Monthly_Balance")
par(mfrow = c(4,2))
suppressWarnings(suppressMessages(hist_num_plots <- lapply(hist_num_cols, function(column_name) {
  ggplot(data, aes(x = !!sym(column_name))) + 
    geom_histogram() + 
    theme_minimal() + 
    labs(title = paste("Histogram of", column_name), x = column_name)
})))

suppressWarnings(suppressMessages(grid.arrange(grobs = hist_num_plots, ncol = 2))) ## display plot

```

<br> <strong> <u>Interpretation:</u></strong>
<br> We can see that because of outliers, the distribution of these columns become very skewed, which is not normal.

<br><br><strong>3.3. Check unique values by each customer</strong>

```{r}
checking_col <- c("Age", "Annual_Income", "Num_Bank_Accounts", "Num_Credit_Card", "Interest_Rate", "Num_of_Loan", "Num_of_Delayed_Payment", "Num_Credit_Inquiries", "Outstanding_Debt", "Total_EMI_per_month")
check_results <- sapply(checking_col, function(x) {
  result <- analyze_unique_values(x)
  return (result)
})

unique_check_df <- data.frame(Columns = checking_col, HasUniqueValuePerCustomer = check_results, row.names = NULL)
unique_check_df

```
<strong> <u>Interpretation:</u></strong>
<br>Excluding the outliers (very large abnormal number), there are some columns only have 1 value per customer. Therefore, it's likely that these are the outliers appearing because of some mistakes in data collecting process. For those columns with FALSE result, after having further inspection, we could see that although these columns have more than 1 value per customer, these values are also duplicated within 1 customer group, except the outliers. Also, these duplicated values are in ordered from top to bottom within that group. Conclusively, we plan to replace the outliers in "TRUE" columns with mode value of each customer and replace the outliers in "FALSE" columns with nearest neighbor value.

<br><br><strong>3.4. Check value for Credit_Utilization_Ratio column</strong>

```{r}
out_of_range <- data %>%
  filter(Credit_Utilization_Ratio < 0 | Credit_Utilization_Ratio > 100)
if(nrow(out_of_range) > 0) {
  print("There are values outside the 0 to 100 range")
} else {
  print("All values are in 0 to 100 range")
}
```

<strong>4. Modify data values</strong>

<strong>4.1. Unique value columns</strong>
<br>For these columns, we clean the outliers with the mode from each group

```{r}
data <- clean_data_with_mode("Annual_Income")
data <- clean_data_with_mode("Interest_Rate")
data <- clean_data_with_mode("Num_of_Loan")
data <- clean_data_with_mode("Outstanding_Debt")

```

<strong>4.2. Non-unique value columns</strong>
<br>For these columns, we clean the outliers with the data from nearest neighbor

```{r}
upper_bound_age <- calculate_bounds("Age")[2]
data <- clean_data_with_neighbor("Age", 0, upper_bound_age)

upper_bound_num_bank_acc <- calculate_bounds("Num_Bank_Accounts")[2]
data <- clean_data_with_neighbor("Num_Bank_Accounts", 0, upper_bound_num_bank_acc)

upper_bound_num_credit_card <- calculate_bounds("Num_Credit_Card")[2]
data <- clean_data_with_neighbor("Num_Credit_Card", 0, upper_bound_num_credit_card)

data <- fill_NA_with_neighbour("Num_of_Delayed_Payment")
upper_bound_num_delay_payment <- calculate_bounds("Num_of_Delayed_Payment")[2]
data <- clean_data_with_neighbor("Num_of_Delayed_Payment", 0, upper_bound_num_delay_payment)

data <- fill_NA_with_neighbour("Num_Credit_Inquiries")
upper_bound_num_credit_inquires <- calculate_bounds("Num_Credit_Inquiries")[2]
data <- clean_data_with_neighbor("Num_Credit_Inquiries", 0, upper_bound_num_credit_inquires)

lower_bound <- calculate_bounds("Total_EMI_per_month")[1]
upper_bound_total_emi <- calculate_bounds("Total_EMI_per_month")[2]
data <- clean_data_with_neighbor("Total_EMI_per_month", 0, 1800) # this column has upper_bound to low, which makes value of some group becomes outlier; therefore, we decide to choose the value of 1800; which is the maximum of valid values among every customer.
```

<strong>4.3. Other columns</strong>

- For column "Monthly_Inhand_Salary", we acknowledge that for each customer, excluding the missing values, the valid values are in groups and in order. Therefore, we decided to fill the missing values based on the nearest neighbor values.
- For column "Delay_from_due_date", we acknowledge that the negative values don't follow any rule or pattern. Therefore, we suppose these negative values are generated from some calculations to calculate the average delay from due date so if it returned negative values, it's possibly indicate that there is no delay from due date anymore. This is why we decided to replace the negative value with 0
- For column "Changed_Credit_Limit", we acknowledge that for each customer, excluding the missing values, the valid values don't include outliers and for some groups, the values are not in groups and not in order. Therefore, we decided to fill the NA with mean of each customer.
- For column "Credit_History_Age", we acknowledge that for each customer, the number of months in lower row is added 1 month from the above row. Therefore, we decided to fill the NA by this approach (+/- 1 month from upper or lower row)
- For columns of "Amount_invested_monthly" and "Monthly_Balance", there are 2 abnormal values "100000" and "3333....", which supposed to be returned error when collecting data. Therefore, for these 2 continuous variables, we decided to replace those abnormal values with the mean of each group

```{r}
data <- fill_NA_with_neighbour("Monthly_Inhand_Salary")
data <- replace_negative_with_0("Delay_from_due_date")
data <- fill_NA_with_mean("Changed_Credit_Limit")
data <- data %>%
  group_by(Customer_ID) %>%
  mutate(Index = row_number()) %>%
  arrange(Customer_ID, Index) %>%
  mutate(First_Valid = min(Index[!is.na(Credit_History_Age)])) %>% 
  mutate(Base = Credit_History_Age[First_Valid] - First_Valid) %>% 
  mutate(Credit_History_Age = if_else(is.na(Credit_History_Age), Index + Base, Credit_History_Age)) %>%
  select(-Index, -First_Valid, -Base) %>%
  ungroup()
data <- replace_with_group_mean("Amount_invested_monthly", 10000)
data <- data %>%
  mutate(Monthly_Balance = ifelse(Monthly_Balance == -333333333333333333333333333, 333333333333333333333333333, Monthly_Balance))
data <- replace_with_group_mean("Monthly_Balance", 333333333333333333333333333)
```

## Categorical variables
<strong>1. Change data type to Factor</strong>

```{r}
categorical_var <- c("Month","Occupation", "Credit_Mix", "Payment_of_Min_Amount", "Payment_Behaviour", "Credit_Score")
data[categorical_var] <- lapply(data[categorical_var], as.factor)
```

<strong>2. Check distribution</strong>

```{r}
# Remove all "and "
data$Type_of_Loan <- gsub("and ", "", data$Type_of_Loan)
# Split by ", "
splitLoans <- strsplit(data$Type_of_Loan, ", ", fixed = TRUE)
# Flatten the list to a vector
loanTypes <- unlist(splitLoans)
# Count the frequency of each text
loanTypeFrequency <- table(loanTypes)
loanTypeFrequencyDF <- as.data.frame(loanTypeFrequency)
colnames(loanTypeFrequencyDF) <- c("LoanType", "Frequency")

type_of_loan_plot <- ggplot(loanTypeFrequencyDF, aes(x = LoanType, y = Frequency, fill = LoanType)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Frequency), vjust = -0.3, color = "black") +
  theme_minimal() +
  labs(title = "Frequency Distribution of Type_of_Loan", x = "Loan Type", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 25, hjust = 1)) +
  scale_fill_brewer(palette = "Spectral")
type_of_loan_plot   
plot_list_1 <- lapply(c("Month", "Occupation", "Credit_Mix", "Payment_of_Min_Amount"), function(col) {
    frequency <- table(data[col])
    frequencyDF <- as.data.frame(frequency)
    names(frequencyDF) <- c("Level", "Frequency")
    return (ggplot(frequencyDF, aes(x = Level, y = Frequency)) +
      geom_bar(stat = "identity", fill = "steelblue") +
      geom_text(aes(label = Frequency), vjust = -0.3, color = "black") + 
      theme_minimal() +
      labs(title = paste("Distribution of", col), x = "Category", y = "Frequency") +
      theme(axis.text.x = element_text(angle = 0, hjust = 1)))
})

plot_list_2 <- lapply(c("Payment_Behaviour", "Credit_Score"), function(col) {
    frequency <- table(data[col])
    frequencyDF <- as.data.frame(frequency)
    names(frequencyDF) <- c("Level", "Frequency")

    return(ggplot(frequencyDF, aes(x = Level, y = Frequency, fill = Level)) +
      geom_bar(stat = "identity") +
      geom_text(aes(label = Frequency), vjust = -0.3, color = "black") +
      theme_minimal() +
      labs(title = paste("Distribution of", col), x = "Category", y = "Frequency") +
      theme(axis.text.x = element_text(angle = 18, hjust = 1)) +
      scale_fill_viridis_d())
})
plot_list <- c(plot_list_1, plot_list_2)
plot_list
```
<br><strong> <u>Interpretation:</u></strong>
<br> We could see that there are 4 variables have invalid values, which are Occupation, Payment_Behaviour, Credit_Mix, and Payment_of_Min_Amount

<br><br><strong>3. Check % of missing/invalid values</strong>

```{r}
missing_val_check_col <- c("Month", "Type_of_Loan", "Occupation", "Credit_Mix" , "Payment_of_Min_Amount", "Payment_Behaviour" , "Credit_Score")
invalid_val <- c("", "", "_______", "_", "NM", "!@9#%8", "")
invalid_val_col <- setNames(as.list(invalid_val), missing_val_check_col)
num_na_list <- sapply(missing_val_check_col, function(x) {
  return (sum(is.na(data[x])))
})
num_invalid_list <- sapply(missing_val_check_col, function(x) {
  return (sum(data[x] == invalid_val_col[x]))
})
percentage_invalid_val_list <- sapply(num_invalid_list, function(x) {
  return((x / nrow(data)) * 100)
})

missing_val_percentage_df <- data.frame(Columns = missing_val_check_col, NumberOfNA = num_na_list, NumberOfInvalidValue = num_invalid_list, PercentageOfInvalidVal = percentage_invalid_val_list, row.names = NULL)
missing_val_percentage_df

```
<strong> <u>Interpretation:</u></strong>
<br> We can see that for Month and Credit_Score variable, there is no missing/invalid value. However for the other variables, there are many missing/invalid values, which have a relatively high percentage. Therefore, we decided to replace these values rather than deleting them.

<br><br><strong>4. Check if excluding invalid value, the values are unique by each customer group</strong>
```{r}
inv_col_check <- c("Occupation", "Credit_Mix" , "Payment_of_Min_Amount", "Payment_Behaviour")
invalid_val_1 <- c("_______", "_", "NM", "!@9#%8")
invalid_val_1_col <- setNames(as.list(invalid_val_1), inv_col_check)

count_unique_value <- function(column_name, bad_value) {
  col_sym <- rlang::sym(column_name)
  df_without_invalid_val <- data[data[column_name] != bad_value,  c("Customer_ID", column_name)]
  count_val <- unique(df_without_invalid_val %>% group_by(Customer_ID, !!(col_sym)))
  return(sum(duplicated(count_val$Customer_ID)))
}

check_inv_result <- sapply(inv_col_check, function(x) {
  result <- count_unique_value(x, invalid_val_1_col[x])
  return (result == 0)
})

unique_val_cat_df <- data.frame(Column = inv_col_check, HasUniqueValuePerCustomer = check_inv_result, row.names = NULL)
unique_val_cat_df
```

<strong><u>Interpretation:</u></strong>
<br> We can see that except Payment_Behavior, the other variables have only 1 value per customer (exclude the invalid value). Also, after further inspecting Payment_Behavior, except the invalid values, the other values are comprehendable. Therefore, we decided to replace these invalid values for these 4 columns with the mode of each customer. For the columns has unique value per customer, the mode of 1 customer is also that unique value.

<br><br><strong>5. Handle missing/invalid values</strong>

<strong>5.1. Handle Type_of_Loan column</strong>

It has been observed that when the values in the Num_of_Loan column are 0 or negative, there still exists some content in the Type_of_Loan column(e.g "Not Specified"). It is considered that a loan count of 0 or negative effectively means there is no loan type. Therefore, the values in the Type_of_Loan column corresponding to 0 or negative values in the Num_of_Loan column should be set to "No_Loan".

```{r}
# Set Type_of_Loan to "No Loan" where Num_of_Loan is 0 or negative
data$Type_of_Loan <- ifelse(data$Num_of_Loan <= 0, "No Loan", data$Type_of_Loan)
```

<strong>5.2. Handle the other columns</strong>
<br> For these columns, we replace the invalid values with mode of each group
```{r}
find_mode_except_bad_value <- function(x, bad_value) {
  ux <- unique(na.omit(x[x != bad_value]))
  ux[which.max(tabulate(match(x, ux)))]
}

data <- data %>%
  group_by(Customer_ID) %>%
  mutate(Mode_Occupation = find_mode_except_bad_value(Occupation, '_______'),
         Occupation = if_else(Occupation == '_______', Mode_Occupation, Occupation),
         Mode_Credit_Mix = find_mode_except_bad_value(Credit_Mix, '_'),
         Credit_Mix = if_else(Credit_Mix == '_', Mode_Credit_Mix, Credit_Mix),
         Mode_Payment_of_Min_Amount = find_mode_except_bad_value(Payment_of_Min_Amount, 'NM'),
         Payment_of_Min_Amount = if_else(Payment_of_Min_Amount == 'NM', Mode_Payment_of_Min_Amount, Payment_of_Min_Amount),
          Mode_Payment_Behaviour = find_mode_except_bad_value(Payment_Behaviour, '!@9#%8'),
         Payment_Behaviour = if_else(Payment_Behaviour == '!@9#%8', Mode_Payment_Behaviour, Payment_Behaviour)) %>%
  select(-c(Mode_Occupation, Mode_Credit_Mix, Mode_Payment_of_Min_Amount, Mode_Payment_Behaviour)) %>%
  ungroup()
```

<strong>6. Assign dummy values to categorical variables</strong>

```{r}
data <- data %>%
  mutate(
    Month_numeric = case_when(
      Month == "January" ~ 1,
      Month == "February" ~ 2,
      Month == "March" ~ 3,
      Month == "April" ~ 4,
      Month == "May" ~ 5,
      Month == "June" ~ 6,
      Month == "July" ~ 7,
      Month == "August" ~ 8,
      TRUE ~ NA_real_
    ),
    Occupation_numeric = case_when(
      Occupation == "Scientist" ~ 1,
      Occupation == "Teacher" ~ 2,
      Occupation == "Engineer" ~ 3,
      Occupation == "Entrepreneur" ~ 4,
      Occupation == "Developer" ~ 5,
      Occupation == "Lawyer" ~ 6,
      Occupation == "Media_Manager" ~ 7,
      Occupation == "Doctor" ~ 8,
      Occupation == "Journalist" ~ 9,
      Occupation == "Manager" ~ 10,
      Occupation == "Accountant" ~ 11,
      Occupation == "Musician" ~ 12,
      Occupation == "Mechanic" ~ 13,
      Occupation == "Writer" ~ 14,
      Occupation == "Architect" ~ 15,
      TRUE ~ NA_real_
    ),
    Credit_Mix_numeric = case_when(
      Credit_Mix == "Bad" ~ 1,
      Credit_Mix == "Standard" ~ 2,
      Credit_Mix == "Good" ~ 3,
      TRUE ~ NA_real_
    ),
    Payment_of_Min_Amount_numeric = case_when(
      Payment_of_Min_Amount == "No" ~ 0,
      Payment_of_Min_Amount == "Yes" ~ 1,
      TRUE ~ NA_real_
    ),
    Payment_Behaviour_numeric = case_when(
      Payment_Behaviour == "High_spent_Large_value_payments" ~ 6,
      Payment_Behaviour == "High_spent_Medium_value_payments" ~ 5,
      Payment_Behaviour == "High_spent_Small_value_payments" ~ 4,
      Payment_Behaviour == "Low_spent_Large_value_payments" ~ 3,
      Payment_Behaviour == "Low_spent_Medium_value_payments" ~ 2,
      Payment_Behaviour == "Low_spent_Small_value_payments" ~ 1,    
      TRUE ~ NA_real_
    ),
    Credit_Score_numeric = case_when(
      Credit_Score == "Poor" ~ 1,
      Credit_Score == "Standard" ~ 2,
      Credit_Score == "Good" ~ 3,
      TRUE ~ NA_real_
    )
  )
```
## Check cleaned dataset
<strong>1. Numerical variables</strong>
```{r}
hist_num_cols <- c("Age", "Num_Bank_Accounts", "Num_Credit_Card", "Num_of_Delayed_Payment", "Num_Credit_Inquiries", "Total_EMI_per_month", "Amount_invested_monthly", "Monthly_Balance")
par(mfrow = c(4,2))
suppressMessages(hist_num_plots <- lapply(hist_num_cols, function(column_name) {
  ggplot(data, aes(x = !!sym(column_name))) + 
    geom_histogram() + 
    theme_minimal() + 
    labs(title = paste("Histogram of", column_name), x = column_name)
}))
suppressMessages(grid.arrange(grobs = hist_num_plots, ncol = 2)) ## display plot
```
<br><strong><u>Conclusion:</u></strong>
<br>We could see that the distribution of numerical variables are improved
<br><strong>2. Categorical variables</strong>
```{r}
# Remove all "and "
data$Type_of_Loan <- gsub("and ", "", data$Type_of_Loan)
# Split by ", "
splitLoans <- strsplit(data$Type_of_Loan, ", ", fixed = TRUE)
# Flatten the list to a vector
loanTypes <- unlist(splitLoans)
# Count the frequency of each text
loanTypeFrequency <- table(loanTypes)
loanTypeFrequencyDF <- as.data.frame(loanTypeFrequency)
colnames(loanTypeFrequencyDF) <- c("LoanType", "Frequency")

type_of_loan_plot <- ggplot(loanTypeFrequencyDF, aes(x = LoanType, y = Frequency, fill = LoanType)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Frequency), vjust = -0.3, color = "black") +
  theme_minimal() +
  labs(title = "Frequency Distribution of Type_of_Loan", x = "Loan Type", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 25, hjust = 1)) +
  scale_fill_brewer(palette = "Spectral")
type_of_loan_plot   
plot_list_1 <- lapply(c("Occupation", "Credit_Mix", "Payment_of_Min_Amount"), function(col) {
    frequency <- table(data[col])
    frequencyDF <- as.data.frame(frequency)
    names(frequencyDF) <- c("Level", "Frequency")
    return (ggplot(frequencyDF, aes(x = Level, y = Frequency)) +
      geom_bar(stat = "identity", fill = "steelblue") +
      geom_text(aes(label = Frequency), vjust = -0.3, color = "black") + 
      theme_minimal() +
      labs(title = paste("Distribution of", col), x = "Category", y = "Frequency") +
      theme(axis.text.x = element_text(angle = 0, hjust = 1)))
})

plot_list_2 <- lapply(c("Payment_Behaviour"), function(col) {
    frequency <- table(data[col])
    frequencyDF <- as.data.frame(frequency)
    names(frequencyDF) <- c("Level", "Frequency")

    return(ggplot(frequencyDF, aes(x = Level, y = Frequency, fill = Level)) +
      geom_bar(stat = "identity") +
      geom_text(aes(label = Frequency), vjust = -0.3, color = "black") +
      theme_minimal() +
      labs(title = paste("Distribution of", col), x = "Category", y = "Frequency") +
      theme(axis.text.x = element_text(angle = 18, hjust = 1)) +
      scale_fill_viridis_d())
})
plot_list <- c(plot_list_1, plot_list_2)
plot_list
```
<br><strong><u>Conclusion:</u></strong>
<br>We could see that there are no more missing/invalid values

# EDA {.tabset .tabset-fade .tabset-pills}
## Heatmap
```{r}

```

## PCA Plot
```{r}
pca.data <- data[, -which(names(data) %in% c("Customer_ID"))] %>% sample_n(1000)
data.att <- select(pca.data, -'Credit_Score')
data.att <- model.matrix(~ . - 1, data = data.att)
credit_score <- factor(pca.data[["Credit_Score"]])
pca.scaled <- prcomp(data.att, scale = TRUE)
credit_score.df <- data.frame(PC1 = pca.scaled$x[, 1], PC2 = pca.scaled$x[, 2], PC3 = pca.scaled$x[, 3], labels = credit_score)
pca.plot <- ggplot(credit_score.df, aes(PC1, PC2, col = labels)) + geom_point() + theme_minimal()
print(pca.plot)
```

# Project Plan {.tabset .tabset-fade .tabset-pills}
<table>
  <tr>
    <th style="border: 1px solid black; padding: 8px">Item</th>
    <th style="border: 1px solid black; padding: 8px">Description</th> 
    <th style="border: 1px solid black; padding: 8px">Due date</th>
    <th style="border: 1px solid black; padding: 8px">Assigned member</th>
  </tr>
  <tr>
    <td style="border: 1px solid black; padding: 8px">Feature engineering</td>
    <td style="border: 1px solid black; padding: 8px">This work is to check again which features need to be transformed to make the better model.</td> 
    <td style="border: 1px solid black; padding: 8px"></td>
    <td style="border: 1px solid black; padding: 8px"></td>
  </tr>
  <tr>
    <td style="border: 1px solid black; padding: 8px">Check the imputation again</td>
    <td style="border: 1px solid black; padding: 8px">This work is to make sure there is no data leakage problem. This step includes decision of where to put the imputation relative to cross validation step</td> 
    <td style="border: 1px solid black; padding: 8px"></td>
    <td style="border: 1px solid black; padding: 8px"></td>
  </tr>
  <tr>
    <td style="border: 1px solid black; padding: 8px">Apply cross validation</td>
    <td style="border: 1px solid black; padding: 8px">To split the dataset into train-test set.</td> 
    <td style="border: 1px solid black; padding: 8px"></td>
    <td style="border: 1px solid black; padding: 8px"></td>
  </tr>
  <tr>
    <td style="border: 1px solid black; padding: 8px">Apply LDA classification method</td>
    <td style="border: 1px solid black; padding: 8px">As our data has three response classes, we chose LDA, which is more intuitive in this situation (Sun, 2024), as one of the modeling techniques. </td>
    <td style="border: 1px solid black; padding: 8px"></td>
    <td style="border: 1px solid black; padding: 8px"></td>
  </tr>
  <tr>
    <td style="border: 1px solid black; padding: 8px">Apply kNN classification method</td>
    <td style="border: 1px solid black; padding: 8px">According to the result of PCA, we can see the decision boundary is non-linear. Since the kNN performs better compared with LDA and Logistic Regression when the decision boundary is highly non-linear (Sun, 2024), we will also use kNN in this project.</td>
    <td style="border: 1px solid black; padding: 8px"></td>
    <td style="border: 1px solid black; padding: 8px"></td>
  </tr>
  <tr>
    <td style="border: 1px solid black; padding: 8px">Apply SVM classification method</td>
    <td style="border: 1px solid black; padding: 8px">In SVM computation, the negative effect of high dimension can be reduced by employing the principle of large margins and kernel functions (Chen, 2004). With the help of kernels, the SVM can also be used for non-linear classification after the late 20th century (Ghosh, Dasgupta & Swetapadma, 2019). So we chose SVM as another classification method. Specifically, since we only have three classes, which is a small number, we will use the one vs one in this project.</td>
    <td style="border: 1px solid black; padding: 8px"></td>
    <td style="border: 1px solid black; padding: 8px"></td>
  </tr>
  <tr>
    <td style="border: 1px solid black; padding: 8px">Apply Decision-tree classification method</td>
    <td style="border: 1px solid black; padding: 8px">Patelli et al. (2023) validated the high flexibility and capability of Random Forest to handle nonlinear data. Shanmugasundar et al. (2021), by comparing linear regression, Random Forest, and AdaBoost regression, demonstrated that Random Forest has significant accuracy and high applicability, making it suitable for large datasets. We will use Random Forest as our fourth classification method.</td>
    <td style="border: 1px solid black; padding: 8px"></td>
    <td style="border: 1px solid black; padding: 8px"></td>
  </tr>
  <tr>
    <td style="border: 1px solid black; padding: 8px">Compare the performance metrics of classification methods</td>
    <td style="border: 1px solid black; padding: 8px">The `confusionMatrix()` function in R simplifies the computation of key classification metrics like accuracy, precision, recall, and F1-score. <br> Of these metrics, the F1-score holds particular importance, offering a comprehensive evaluation of model performance by considering both precision and recall alongside accuracy. This holistic assessment helps gauge the model's predictive capacity effectively.
<br>Moreover, delving into ROC curve analysis and examining the AUC value provides deeper insights into the model's classification prowess. This analysis aids in identifying optimal thresholds to balance recall and false positive rate, thereby assessing the model's overall effectiveness. Additionally, ROC curve analysis facilitates model comparison, assisting in the selection of the most suitable model for specific requirements.
</td>
    <td style="border: 1px solid black; padding: 8px"></td>
    <td style="border: 1px solid black; padding: 8px"></td>
  </tr>
  <tr>
    <td style="border: 1px solid black; padding: 8px">Write final report</td>
    <td style="border: 1px solid black; padding: 8px"></td>
    <td style="border: 1px solid black; padding: 8px"></td>
    <td style="border: 1px solid black; padding: 8px"></td>
  </tr>
  <tr>
    <td style="border: 1px solid black; padding: 8px">Prepare presentation slides</td>
    <td style="border: 1px solid black; padding: 8px"></td>
    <td style="border: 1px solid black; padding: 8px"></td>
    <td style="border: 1px solid black; padding: 8px"></td>
  </tr>
  <tr>
    <td style="border: 1px solid black; padding: 8px">Record presentation</td>
    <td style="border: 1px solid black; padding: 8px"></td>
    <td style="border: 1px solid black; padding: 8px"></td>
    <td style="border: 1px solid black; padding: 8px"></td>
  </tr>
</table>