---
title: "Group Project"
author: "Group 38"
date: "2024-04-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Load libraries
```{r}
library('tidyverse')
library('readr')
library('dplyr')
library('ggplot2')
library('broom')
library('tinytex')
library('caret')
```

## Load dataset

```{r}
data <- as.data.frame(read.csv("credit_train.csv", header = TRUE))
head(data)
```
## Check duplicate rows
```{r}
sum(duplicated(data))
```

## Remove unnecessary columns
```{r}
data <- data[ , -which(names(data) %in% c("ID", "Name", "SSN"))]
```

## Change value for Credit_History_Age column

```{r}
data["Credit_History_Age"] = lapply(data["Credit_History_Age"], function(x) {
    years <- as.integer(sub("\\Years.*", "", x))
    months <- as.integer(sub(".*\\d+\\s+Years\\s+and\\s+(\\d+)\\s+Months.*", "\\1", x))
    return(years*12 + months)
})
```


# Numerical var
## Delete "_"
```{r}
for(i in c("Age","Annual_Income", "Num_of_Loan", "Num_of_Delayed_Payment", 
                    "Changed_Credit_Limit", "Outstanding_Debt", "Amount_invested_monthly", "Monthly_Balance")) {
  data[[i]] <- gsub("_", "", data[[i]])
  
  data[[i]] <- as.numeric(data[[i]])
}

find_most <- function(x) {
  ux <- unique(na.omit(x))
  ux[which.max(tabulate(match(x, ux)))]
}
```
```{r}
colSums(is.na(data))
```
## Calculate the percentage of missing values for each numerical column
```{r}
missing_values <- colSums(is.na(data[c("Age", "Annual_Income", "Monthly_Inhand_Salary", "Num_Bank_Accounts", 
                                       "Num_Credit_Card", "Interest_Rate", "Num_of_Loan", 
                                       "Delay_from_due_date", "Num_of_Delayed_Payment", 
                                       "Changed_Credit_Limit", "Num_Credit_Inquiries", 
                                       "Outstanding_Debt", "Credit_Utilization_Ratio", 
                                       "Credit_History_Age", "Total_EMI_per_month", 
                                       "Amount_invested_monthly", "Monthly_Balance")]))


total_values <- nrow(data)

missing_values_percent <- missing_values / total_values * 100

print(missing_values_percent)
```


## Age
```{r}
# This is a function to find the bound of the column by using IQR
calculate_bounds <- function(column_name) {
  Q1 <- quantile(data[[column_name]], 0.25, na.rm = TRUE)
  Q3 <- quantile(data[[column_name]], 0.75, na.rm = TRUE)
  IQR <- Q3 - Q1
  lower_bound <- Q1 - 1.5 * IQR
  upper_bound <- Q3 + 1.5 * IQR
  return(c(lower_bound, upper_bound))
}
```
```{r}
# This is a function to check if each group has a unique value
analyze_unique_values <- function(income_column) {
  lower_bound <- calculate_bounds(income_column)[1]
  upper_bound <- calculate_bounds(income_column)[2]
  
  filtered_data <- data %>%
    filter(!!sym(income_column) >= lower_bound & !!sym(income_column) <= upper_bound)
  
  result <- filtered_data %>%
    group_by(Customer_ID) %>%
    summarise(Unique_Value_Count = n_distinct(!!sym(income_column)),
              Only_One_Unique_Value = Unique_Value_Count == 1) %>%
    ungroup()

  if(all(result$Only_One_Unique_Value)) {
    cat("All groups have only one unique value.\n")
  } else {
    cat("These groups have different values:\n")
    print(result %>% filter(Only_One_Unique_Value == FALSE))
  }
}
```

```{r}
analyze_unique_values("Age")
```
```{r}
# This is a function to change outliers with neighbor
clean_data_with_neighbor <- function(column_name, lower_bound, upper_bound) {
  column_sym <- sym(column_name)
  result <- data %>%
    group_by(Customer_ID) %>%
    mutate(
      !!column_sym := ifelse(
        .data[[column_name]] < lower_bound | .data[[column_name]] > upper_bound, NA, .data[[column_name]]
      )
    ) %>%
    fill(!!column_sym, .direction = "up") %>%
    fill(!!column_sym, .direction = "down") %>%
    ungroup()
  
  return(result)
}
```
```{r}
lower_bound <- calculate_bounds("Age")[1]
upper_bound <- calculate_bounds("Age")[2]

data <- clean_data_with_neighbor("Age", 0, upper_bound)
```


## Annual_Income
```{r}
analyze_unique_values("Annual_Income")
```
```{r}
# This is a function that can find the most common value in every column
find_mode <- function(x) {
  ux <- unique(na.omit(x))
  ux[which.max(tabulate(match(x, ux)))]
}
```
```{r}
# This is a function to change outliers with mode
clean_data_with_mode <- function(column_name) {
  col_sym <- rlang::sym(column_name) 
  result <- data %>%
    group_by(Customer_ID) %>%
    mutate(Mode = find_mode(!!col_sym),
           !!col_sym := Mode) %>% 
    select(-Mode) %>%
    ungroup()
  
  return(result)
}
```
```{r}
data <- clean_data_with_mode("Annual_Income")
```
## Monthly_Inhand_Salary

```{r}
# This is a function to fill NA with neighbor
fill_NA_with_neighbour <- function(column_name) {
  result <- data %>%
    group_by(Customer_ID) %>%
    fill(!!rlang::sym(column_name), .direction = "up") %>%
    fill(!!rlang::sym(column_name), .direction = "down") %>%
    ungroup()
  
  return(result)
}
```
```{r}
data <- fill_NA_with_neighbour("Monthly_Inhand_Salary")
```
## Num_Bank_Accounts
```{r}
analyze_unique_values("Num_Bank_Accounts")
```
```{r}
lower_bound <- calculate_bounds("Num_Bank_Accounts")[1]
upper_bound <- calculate_bounds("Num_Bank_Accounts")[2]

data <- clean_data_with_neighbor("Num_Bank_Accounts", 0, upper_bound)
```


## Num_Credit_Card
```{r}
analyze_unique_values("Num_Credit_Card")
```
```{r}
lower_bound <- calculate_bounds("Num_Credit_Card")[1]
upper_bound <- calculate_bounds("Num_Credit_Card")[2]

data <- clean_data_with_neighbor("Num_Credit_Card", 0, upper_bound)
```


## Interest_Rate
```{r}
analyze_unique_values("Interest_Rate")
```
```{r}
data <- clean_data_with_mode("Interest_Rate")
```


## Num_of_Loan
```{r}
analyze_unique_values("Num_of_Loan")
```
```{r}
data <- clean_data_with_mode("Num_of_Loan")
```


## Delay_from_due_date
```{r}
# This is a function to replace negative value with 0
replace_negative_with_0 <- function(column_name) {
  result <- data %>%
    mutate(!!rlang::sym(column_name) := ifelse(!!rlang::sym(column_name) < 0, 0, !!rlang::sym(column_name)))
  
  return(result)
}
```
```{r}
data <- replace_negative_with_0("Delay_from_due_date")
```

## Num_of_Delayed_Payment
```{r}
data <- fill_NA_with_neighbour("Num_of_Delayed_Payment")
lower_bound <- calculate_bounds("Num_of_Delayed_Payment")[1]
upper_bound <- calculate_bounds("Num_of_Delayed_Payment")[2]

data <- clean_data_with_neighbor("Num_of_Delayed_Payment", 0, upper_bound)
```


## Changed_Credit_Limit
```{r}
data <- replace_negative_with_0("Changed_Credit_Limit")
```
```{r}
# This is a function to fill NA with mean
fill_NA_with_mean <- function(column_name) {
  col_sym <- rlang::sym(column_name)
  result <- data %>%
    group_by(Customer_ID) %>%
    mutate(
      Mean = round(mean({{col_sym}}[!is.na({{col_sym}})], na.rm = TRUE), 2),
      {{col_sym}} := if_else(is.na({{col_sym}}), Mean, {{col_sym}})
    ) %>%
    select(-Mean) %>%
    ungroup()
  return(result)
}
```
```{r}
data <- fill_NA_with_mean("Changed_Credit_Limit")
```


## Num_Credit_Inquiries
```{r}
analyze_unique_values("Num_Credit_Inquiries")
```
```{r}
data <- fill_NA_with_neighbour("Num_Credit_Inquiries")
lower_bound <- calculate_bounds("Num_Credit_Inquiries")[1]
upper_bound <- calculate_bounds("Num_Credit_Inquiries")[2]

data <- clean_data_with_neighbor("Num_Credit_Inquiries", 0, upper_bound)
```


## Outstanding_Debt
```{r}
analyze_unique_values("Outstanding_Debt")
```
```{r}
data <- clean_data_with_mode("Outstanding_Debt")
```


## Credit_Utilization_Ratio
```{r}
out_of_range <- data %>%
  filter(Credit_Utilization_Ratio < 0 | Credit_Utilization_Ratio > 100)
print(out_of_range)
if(nrow(out_of_range) > 0) {
  print("There are values outside the 0 to 100 range")
} else {
  print("All values are in 0 to 100 range")
}
```


## Credit_History_Age
```{r}
data <- data %>%
  group_by(Customer_ID) %>%
  mutate(Index = row_number()) %>%
  arrange(Customer_ID, Index) %>%
  mutate(First_Valid = min(Index[!is.na(Credit_History_Age)])) %>% 
  mutate(Base = Credit_History_Age[First_Valid] - First_Valid) %>% 
  mutate(Credit_History_Age = if_else(is.na(Credit_History_Age), Index + Base, Credit_History_Age)) %>%
  select(-Index, -First_Valid, -Base) %>%
  ungroup()
```


## Total_EMI_per_month

```{r}
analyze_unique_values("Total_EMI_per_month")
```
```{r}
lower_bound <- calculate_bounds("Total_EMI_per_month")[1]
upper_bound <- calculate_bounds("Total_EMI_per_month")[2]

data <- clean_data_with_neighbor("Total_EMI_per_month", 0, 1800)
```


## Amount_invested_monthly
```{r}
unique(data$Amount_invested_monthly)
```
```{r}
# This is a function to replace the outliers and NA with mean
replace_with_group_mean <- function(column_name, value_to_replace) {
  col_sym <- sym(column_name)
  result <- data %>%
    group_by(Customer_ID) %>%
    mutate(
      Group_Mean = mean({{col_sym}}[{{col_sym}} != value_to_replace & !is.na({{col_sym}})], na.rm = TRUE),
      {{col_sym}} := if_else({{col_sym}} == value_to_replace | is.na({{col_sym}}), Group_Mean, {{col_sym}})
    ) %>%
    select(-Group_Mean) %>%
    ungroup()
  return(result)
}
```
```{r}
data <- replace_with_group_mean("Amount_invested_monthly", 10000)
```


## Monthly_Balance
```{r}
unique(data$Monthly_Balance)
```
```{r}
data <- data %>%
  mutate(Monthly_Balance = ifelse(Monthly_Balance == -333333333333333333333333333, 333333333333333333333333333, Monthly_Balance))
```
```{r}
data <- replace_with_group_mean("Monthly_Balance", 333333333333333333333333333)
```
```{r}
colSums(is.na(data))
```
```{r}
data
```

## View the distribution of each numerical column
```{r}
columns <- c("Age", "Annual_Income", "Monthly_Inhand_Salary", "Num_Bank_Accounts", 
             "Num_Credit_Card", "Interest_Rate", "Num_of_Loan", 
             "Delay_from_due_date", "Num_of_Delayed_Payment", 
             "Changed_Credit_Limit", "Num_Credit_Inquiries", 
             "Outstanding_Debt", "Credit_Utilization_Ratio", 
             "Credit_History_Age", "Total_EMI_per_month", 
             "Amount_invested_monthly", "Monthly_Balance")

par(mfrow = c(3, 6)) # Adjust the layout based on the number of plots you want per row and column

for (column_name in columns) {
  graph <- ggplot(data, aes(y = !!sym(column_name))) + 
    geom_boxplot(fill = "lightblue", color = "black") + 
    theme_minimal() + 
    labs(title = paste("Boxplot of", column_name), y = column_name) + 
    theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.title = element_text(hjust = 0.5))
  
  print(graph)
}

```

# Categorical var
## Change data types
```{r}
categorical_var <- c("Month","Occupation", "Credit_Mix", "Payment_of_Min_Amount", "Payment_Behaviour", "Credit_Score")
data[categorical_var] <- lapply(data[categorical_var], as.factor)
str(data)
```
# 1. Type_of_Loan

## 1.1 Count NA and empty strings

```{r}
# Count NA and empty strings in the Type_of_Loan column
num_na <- sum(is.na(data$Type_of_Loan))
num_blank <- sum(data$Type_of_Loan == "")
print(paste("Number of missing values:", num_na))
print(paste("Number of blank:", num_blank))

# Calculate the percentage of blank
percentage_blank <- (num_blank / nrow(data)) * 100
print(paste("Percentage of blank:", percentage_blank, "%"))

```
## 1.2 handle with empty strings

```{r}
# Use conditional indexing to get corresponding Type_of_Loan values
non_loan <- data$Type_of_Loan[data$Num_of_Loan <= 0]
head(non_loan)
```

It has been observed that when the values in the Num_of_Loan column are 0 or negative, there still exists some content in the Type_of_Loan column(e.g "Not Specified"). It is considered that a loan count of 0 or negative effectively means there is no loan type. Therefore, the values in the Type_of_Loan column corresponding to 0 or negative values in the Num_of_Loan column should be set to "No_Loan".

```{r}
# Set Type_of_Loan to "Non Loan" where Num_of_Loan is 0 or negative
data$Type_of_Loan <- ifelse(data$Num_of_Loan <= 0, "No Loan", data$Type_of_Loan)
```

## 1.3 distribution of each type

```{r}
# Remove all "and "
data$Type_of_Loan <- gsub("and ", "", data$Type_of_Loan)

# Split by ", "
splitLoans <- strsplit(data$Type_of_Loan, ", ", fixed = TRUE)

# Flatten the list to a vector
loanTypes <- unlist(splitLoans)

# Count the frequency of each text
loanTypeFrequency <- table(loanTypes)

loanTypeFrequencyDF <- as.data.frame(loanTypeFrequency)
colnames(loanTypeFrequencyDF) <- c("LoanType", "Frequency")

library(ggplot2)
ggplot(loanTypeFrequencyDF, aes(x = LoanType, y = Frequency, fill = LoanType)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Frequency), vjust = -0.3, color = "black") +
  theme_minimal() +
  labs(title = "Frequency Distribution of Type_of_Loan", x = "Loan Type", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 25, hjust = 1)) +
  scale_fill_brewer(palette = "Spectral")

```

# 2. Credit_Mix

## 2.1 Count NA and "_" strings

```{r}
levels_of_Credit_Mix <- levels(data$Credit_Mix)
print(levels_of_Credit_Mix)
```


```{r}
# rename "_" with "Unknown"
#levels(data$Credit_Mix) <- c("Unknown", "Bad", "Good", "Standard")

# Count NA and "_" strings in the Credit_Mix column
num_na <- sum(is.na(data$Credit_Mix))
num_unknown <- sum(data$Credit_Mix == "_")
print(paste("Number of missing values:", num_na))
print(paste("Number of unknown:", num_unknown))

# Calculate the percentage of none
percentage_unknown <- (num_unknown / nrow(data)) * 100
print(paste("Percentage of unknown:", percentage_unknown, "%"))
```

## 2.2 distribution of Credit_Mix

```{r}
frequency_of_Credit_Mix <- table(data$Credit_Mix)
frequencyDF <- as.data.frame(frequency_of_Credit_Mix)
names(frequencyDF) <- c("Level", "Frequency")

ggplot(frequencyDF, aes(x = Level, y = Frequency)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = Frequency), vjust = -0.3, color = "black") + 
  theme_minimal() +
  labs(title = "Distribution of Credit_Mix", x = "Category", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 0, hjust = 1))
```

# 3. Payment_of_Min_Amount

## 3.1 Count NA and NM

```{r}
levels_of_Payment_of_Min_Amount <- levels(data$Payment_of_Min_Amount)
print(levels_of_Payment_of_Min_Amount)
```


```{r}
# rename "NM" with "Unknown"
#levels(data$Payment_of_Min_Amount) <- c("Unknown", "No", "Yes")

# Count NA and NM in the Payment_of_Min_Amount column
num_na <- sum(is.na(data$Payment_of_Min_Amount))
num_unknown <- sum(data$Payment_of_Min_Amount == "NM")
print(paste("Number of missing values:", num_na))
print(paste("Number of unknown:", num_unknown))

# Calculate the percentage of none
percentage_unknown <- (num_unknown / nrow(data)) * 100
print(paste("Percentage of unknown:", percentage_unknown, "%"))
```


## 3.3 distribution of Payment_of_Min_Amount

```{r}
frequency_of_Payment_of_Min_Amount <- table(data$Payment_of_Min_Amount)
frequencyDF <- as.data.frame(frequency_of_Payment_of_Min_Amount)
names(frequencyDF) <- c("Level", "Frequency")

ggplot(frequencyDF, aes(x = Level, y = Frequency)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = Frequency), vjust = -0.3, color = "black") + 
  theme_minimal() +
  labs(title = "Distribution of Payment_of_Min_Amount", x = "Category", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 0, hjust = 1))
```


# 4. Payment_Behaviour

## 4.1 Count NA and "!@9#%8"

```{r}
levels_of_Payment_Behaviour <- levels(data$Payment_Behaviour)
print(levels_of_Payment_Behaviour)
```

```{r}
# rename "!@9#%8" with "Unknown"
#levels(data$Payment_Behaviour) <- c("Unknown", "High_Large", "High_Medium", "High_Small", "Low_Large", "Low_Medium", "Low_Small")

# Count NA and "!@9#%8" in the Payment_Behaviour column
num_na <- sum(is.na(data$Payment_Behaviour))
num_unknown <- sum(data$Payment_Behaviour == "!@9#%8")
print(paste("Number of missing values:", num_na))
print(paste("Number of unknown:", num_unknown))

# Calculate the percentage of unknown
percentage_unknown <- (num_unknown / nrow(data)) * 100
print(paste("Percentage of unknown:", percentage_unknown, "%"))
```


## 4.2 distribution of Payment_Behaviour

```{r}
frequency_of_Payment_Behaviour <- table(data$Payment_Behaviour)
frequencyDF <- as.data.frame(frequency_of_Payment_Behaviour)
names(frequencyDF) <- c("Level", "Frequency")

ggplot(frequencyDF, aes(x = Level, y = Frequency, fill = Level)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Frequency), vjust = -0.3, color = "black") +
  theme_minimal() +
  labs(title = "Distribution of Payment_Behaviour", x = "Category", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 18, hjust = 1)) +
  scale_fill_viridis_d()
```


# 5. Credit_Score

## 5.1 Count missing value

```{r}
data$Credit_Score <- factor(data$Credit_Score)
levels_of_Credit_Score <- levels(data$Credit_Score)
print(levels_of_Credit_Score)

# Count missing value
num_na <- sum(is.na(data$Payment_Behaviour))
print(paste("Number of missing values:", num_na))

```


## 5.2 distribution of Credit_Score

```{r}
frequency_of_Credit_Score <- table(data$Credit_Score)
frequencyDF <- as.data.frame(frequency_of_Credit_Score)
names(frequencyDF) <- c("Level", "Frequency")

ggplot(frequencyDF, aes(x = Level, y = Frequency, fill = Level)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Frequency), vjust = -0.3, color = "black") +
  theme_minimal() +
  labs(title = "Distribution of Credit_Score", x = "Category", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 18, hjust = 1)) +
  scale_fill_viridis_d()
```


```{r}
count_occupation <- unique(data[data$Occupation != '_______', which(names(data) %in% c("Customer_ID", "Occupation"))] %>% group_by(Customer_ID, Occupation))
sum(duplicated(count_occupation$Customer_ID))
```

```{r}
count_payment_of_min_amount <- unique(data[data$Payment_of_Min_Amount != 'NM', which(names(data) %in% c("Customer_ID", "Payment_of_Min_Amount"))] %>% group_by(Customer_ID, Payment_of_Min_Amount))
sum(duplicated(count_payment_of_min_amount$Customer_ID))
```
```{r}
count_credit_mix <- unique(data[data$Credit_Mix != '_', which(names(data) %in% c("Customer_ID", "Credit_Mix"))] %>% group_by(Customer_ID, Credit_Mix))
sum(duplicated(count_credit_mix$Customer_ID))
```

```{r}
count_payment_behavior <- data[data$Payment_Behaviour != '!@9#%8', which(names(data) %in% c("Customer_ID", "Payment_Behaviour"))] %>% group_by(Customer_ID, Payment_Behaviour) %>% summarise(total_count = n(), .groups = "drop")
sum(duplicated(count_payment_behavior$Customer_ID))
```
```{r}
find_mode_except_bad_value <- function(x, bad_value) {
  ux <- unique(na.omit(x[x != bad_value]))
  ux[which.max(tabulate(match(x, ux)))]
}

data <- data %>%
  group_by(Customer_ID) %>%
  mutate(Mode_Occupation = find_mode_except_bad_value(Occupation, '_______'),
         Occupation = if_else(Occupation == '_______', Mode_Occupation, Occupation),
         Mode_Credit_Mix = find_mode_except_bad_value(Credit_Mix, '_'),
         Credit_Mix = if_else(Credit_Mix == '_', Mode_Credit_Mix, Credit_Mix),
         Mode_Payment_of_Min_Amount = find_mode_except_bad_value(Payment_of_Min_Amount, 'NM'),
         Payment_of_Min_Amount = if_else(Payment_of_Min_Amount == 'NM', Mode_Payment_of_Min_Amount, Payment_of_Min_Amount),
          Mode_Payment_Behaviour = find_mode_except_bad_value(Payment_Behaviour, '!@9#%8'),
         Payment_Behaviour = if_else(Payment_Behaviour == '!@9#%8', Mode_Payment_Behaviour, Payment_Behaviour)) %>%
  select(-c(Mode_Occupation, Mode_Credit_Mix, Mode_Payment_of_Min_Amount, Mode_Payment_Behaviour)) %>%
  ungroup()
```

```{r}
nrow(data[data$Credit_Mix == '_',])
nrow(data[data$Occupation == '_______',])
nrow(data[data$Payment_of_Min_Amount == 'NM',])
nrow(data[data$Payment_Behaviour == '!@9#%8',])
```

```{r}
sum(is.null(data$Month))
data %>% group_by(Month) %>% tally()
```


# 3 Factorize categorical types

```{r}
data <- data %>%
  mutate(
    Credit_Mix_numeric = case_when(
      Credit_Mix == "Bad" ~ 1,
      Credit_Mix == "Standard" ~ 2,
      Credit_Mix == "Good" ~ 3,
      TRUE ~ NA_real_
    ),
    Payment_of_Min_Amount_numeric = case_when(
      Payment_of_Min_Amount == "No" ~ 0,
      Payment_of_Min_Amount == "Yes" ~ 1,
      TRUE ~ NA_real_
    ),
    Payment_Behaviour_numeric = case_when(
      Payment_Behaviour == "High_spent_Large_value_payments" ~ 6,
      Payment_Behaviour == "High_spent_Medium_value_payments" ~ 5,
      Payment_Behaviour == "High_spent_Small_value_payments" ~ 4,
      Payment_Behaviour == "Low_spent_Large_value_payments" ~ 3,
      Payment_Behaviour == "Low_spent_Medium_value_payments" ~ 2,
      Payment_Behaviour == "Low_spent_Small_value_payments" ~ 1,    
      TRUE ~ NA_real_
    ),
    Credit_Score_numeric = case_when(
      Credit_Score == "Poor" ~ 1,
      Credit_Score == "Standard" ~ 2,
      Credit_Score == "Good" ~ 3,
      TRUE ~ NA_real_
    )
  )

head(data)

```

