---
title: 5003_datacleaning"
author: "Wanxin Bao"
date: "2024-04-06"
output:
   html_document:
    fig_caption: yes 
    number_sections: yes
    self_contained: yes
    theme: flatly
    toc: true
    toc_depth: 2
    toc_float: true
    code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Preparation and data load

```{r, include=FALSE}
library(ggplot2)
library(caret)
library(tidyr)
```

```{r}
data <- read.csv("C:/Users/apple/Desktop/train.csv")
head(data)
```

# 1. Type_of_Loan

## 1.1 Count NA and empty strings

```{r}
# Count NA and empty strings in the Type_of_Loan column
num_na <- sum(is.na(data$Type_of_Loan))
num_blank <- sum(data$Type_of_Loan == "")
print(paste("Number of missing values:", num_na))
print(paste("Number of blank:", num_blank))

# Calculate the percentage of blank
percentage_blank <- (num_blank / nrow(data)) * 100
print(paste("Percentage of blank:", percentage_blank, "%"))

```


## 1.2 handle with empty strings

```{r}
# Use conditional indexing to get corresponding Type_of_Loan values
non_loan <- data$Type_of_Loan[data$Num_of_Loan <= 0]
head(non_loan)
```

It has been observed that when the values in the Num_of_Loan column are 0 or negative, there still exists some content in the Type_of_Loan column(e.g "Not Specified"). It is considered that a loan count of 0 or negative effectively means there is no loan type. Therefore, the values in the Type_of_Loan column corresponding to 0 or negative values in the Num_of_Loan column should be set to "Non_Loan".

```{r}
# Set Type_of_Loan to "Non Loan" where Num_of_Loan is 0 or negative
data$Type_of_Loan <- ifelse(data$Num_of_Loan <= 0, "Non Loan", data$Type_of_Loan)
```


## 1.3 distribution of each type

```{r}
# Remove all "and "
data$Type_of_Loan <- gsub("and ", "", data$Type_of_Loan)

# Split by ", "
splitLoans <- strsplit(data$Type_of_Loan, ", ", fixed = TRUE)

# Flatten the list to a vector
loanTypes <- unlist(splitLoans)

# Count the frequency of each text
loanTypeFrequency <- table(loanTypes)

loanTypeFrequencyDF <- as.data.frame(loanTypeFrequency)
colnames(loanTypeFrequencyDF) <- c("LoanType", "Frequency")

library(ggplot2)
ggplot(loanTypeFrequencyDF, aes(x = LoanType, y = Frequency, fill = LoanType)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Frequency), vjust = -0.3, color = "black") +
  theme_minimal() +
  labs(title = "Frequency Distribution of Type_of_Loan", x = "Loan Type", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 25, hjust = 1)) +
  scale_fill_brewer(palette = "Spectral")

```



# 2. Credit_Mix

## 2.1 Count NA and "_" strings

```{r}
data$Credit_Mix <- factor(data$Credit_Mix)
levels_of_Credit_Mix <- levels(data$Credit_Mix)
print(levels_of_Credit_Mix)
```


```{r}
# rename "_" with "Unknown"
levels(data$Credit_Mix) <- c("Unknown", "Bad", "Good", "Standard")

# Count NA and "_" strings in the Credit_Mix column
num_na <- sum(is.na(data$Credit_Mix))
num_unknown <- sum(data$Credit_Mix == "Unknown")
print(paste("Number of missing values:", num_na))
print(paste("Number of unknown:", num_unknown))

# Calculate the percentage of none
percentage_unknown <- (num_unknown / nrow(data)) * 100
print(paste("Percentage of unknown:", percentage_unknown, "%"))
```

## 2.2 distribution of Credit_Mix

```{r}
frequency_of_Credit_Mix <- table(data$Credit_Mix)
frequencyDF <- as.data.frame(frequency_of_Credit_Mix)
names(frequencyDF) <- c("Level", "Frequency")

ggplot(frequencyDF, aes(x = Level, y = Frequency)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = Frequency), vjust = -0.3, color = "black") + 
  theme_minimal() +
  labs(title = "Distribution of Credit_Mix", x = "Category", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 0, hjust = 1))
```


# 3. Payment_of_Min_Amount

## 3.1 Count NA and NM

```{r}
data$Payment_of_Min_Amount <- factor(data$Payment_of_Min_Amount)
levels_of_Payment_of_Min_Amount <- levels(data$Payment_of_Min_Amount)
print(levels_of_Payment_of_Min_Amount)
```


```{r}
# rename "NM" with "Unknown"
levels(data$Payment_of_Min_Amount) <- c("Unknown", "No", "Yes")

# Count NA and NM in the Payment_of_Min_Amount column
num_na <- sum(is.na(data$Payment_of_Min_Amount))
num_unknown <- sum(data$Payment_of_Min_Amount == "Unknown")
print(paste("Number of missing values:", num_na))
print(paste("Number of unknown:", num_unknown))

# Calculate the percentage of none
percentage_unknown <- (num_unknown / nrow(data)) * 100
print(paste("Percentage of unknown:", percentage_unknown, "%"))
```


## 3.3 distribution of Payment_of_Min_Amount

```{r}
frequency_of_Payment_of_Min_Amount <- table(data$Payment_of_Min_Amount)
frequencyDF <- as.data.frame(frequency_of_Payment_of_Min_Amount)
names(frequencyDF) <- c("Level", "Frequency")

ggplot(frequencyDF, aes(x = Level, y = Frequency)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = Frequency), vjust = -0.3, color = "black") + 
  theme_minimal() +
  labs(title = "Distribution of Payment_of_Min_Amount", x = "Category", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 0, hjust = 1))
```


# 4. Payment_Behaviour

## 4.1 Count NA and "!@9#%8"

```{r}
data$Payment_Behaviour <- factor(data$Payment_Behaviour)
levels_of_Payment_Behaviour <- levels(data$Payment_Behaviour)
print(levels_of_Payment_Behaviour)
```

```{r}
# rename "!@9#%8" with "Unknown"
levels(data$Payment_Behaviour) <- c("Unknown", "High_Large", "High_Medium", "High_Small", "Low_Large", "Low_Medium", "Low_Small")

# Count NA and "!@9#%8" in the Payment_Behaviour column
num_na <- sum(is.na(data$Payment_Behaviour))
num_unknown <- sum(data$Payment_Behaviour == "Unknown")
print(paste("Number of missing values:", num_na))
print(paste("Number of unknown:", num_unknown))

# Calculate the percentage of unknown
percentage_unknown <- (num_unknown / nrow(data)) * 100
print(paste("Percentage of unknown:", percentage_unknown, "%"))
```


## 4.2 distribution of Payment_Behaviour

```{r}
frequency_of_Payment_Behaviour <- table(data$Payment_Behaviour)
frequencyDF <- as.data.frame(frequency_of_Payment_Behaviour)
names(frequencyDF) <- c("Level", "Frequency")

ggplot(frequencyDF, aes(x = Level, y = Frequency, fill = Level)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Frequency), vjust = -0.3, color = "black") +
  theme_minimal() +
  labs(title = "Distribution of Payment_Behaviour", x = "Category", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 18, hjust = 1)) +
  scale_fill_viridis_d()
```


# 5. Credit_Score

## 5.1 Count missing value

```{r}
data$Credit_Score <- factor(data$Credit_Score)
levels_of_Credit_Score <- levels(data$Credit_Score)
print(levels_of_Credit_Score)

# Count missing value
num_na <- sum(is.na(data$Payment_Behaviour))
print(paste("Number of missing values:", num_na))

```


## 5.2 distribution of Credit_Score

```{r}
frequency_of_Credit_Score <- table(data$Credit_Score)
frequencyDF <- as.data.frame(frequency_of_Credit_Score)
names(frequencyDF) <- c("Level", "Frequency")

ggplot(frequencyDF, aes(x = Level, y = Frequency, fill = Level)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Frequency), vjust = -0.3, color = "black") +
  theme_minimal() +
  labs(title = "Distribution of Credit_Score", x = "Category", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 18, hjust = 1)) +
  scale_fill_viridis_d()
```


# 3 Factorize categorical types

```{r}
library(dplyr)

data <- data %>%
  mutate(
    Credit_Mix_numeric = case_when(
      Credit_Mix == "Bad" ~ 1,
      Credit_Mix == "Standard" ~ 2,
      Credit_Mix == "Good" ~ 3,
      TRUE ~ NA_real_
    ),
    Payment_of_Min_Amount_numeric = case_when(
      Payment_of_Min_Amount == "No" ~ 0,
      Payment_of_Min_Amount == "Yes" ~ 1,
      TRUE ~ NA_real_
    ),
    Payment_Behaviour_numeric = case_when(
      Payment_Behaviour == "High_Large" ~ 6,
      Payment_Behaviour == "High_Medium" ~ 5,
      Payment_Behaviour == "High_Small" ~ 4,
      Payment_Behaviour == "Low_Large" ~ 3,
      Payment_Behaviour == "Low_Medium" ~ 2,
      Payment_Behaviour == "Low_Small" ~ 1,    
      TRUE ~ NA_real_
    ),
    Credit_Score_numeric = case_when(
      Credit_Score == "Poor" ~ 1,
      Credit_Score == "Standard" ~ 2,
      Credit_Score == "Good" ~ 3,
      TRUE ~ NA_real_
    )
  )


head(data)
```

