---
title: "Group Project"
author: "Group 38"
date: "2024-04-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Load libraries
```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(caret)
```

## Load dataset

```{r}
data <- as.data.frame(read.csv("credit_train.csv", header = TRUE))
head(data)
```
## Check duplicate rows
```{r}
sum(duplicated(data))
```

## Remove unnecessary columns
```{r}
data <- data[ , -which(names(data) %in% c("ID", "Name", "SSN"))]
```

## Change value for Credit_History_Age column

```{r}
data["Credit_History_Age"] = lapply(data["Credit_History_Age"], function(x) {
    years <- as.integer(sub("\\Years.*", "", x))
    months <- as.integer(sub(".*\\d+\\s+Years\\s+and\\s+(\\d+)\\s+Months.*", "\\1", x))
    return(years*12 + months)
})
```

## Check data values of target variable
```{r}
is.null(data$Credit_Score)
sum(is.na(data$Credit_Score) == TRUE)
```
# 1. Type_of_Loan

## 1.1 Count NA and empty strings

```{r}
# Count NA and empty strings in the Type_of_Loan column
num_na <- sum(is.na(data$Type_of_Loan))
num_blank <- sum(data$Type_of_Loan == "")
print(paste("Number of missing values:", num_na))
print(paste("Number of blank:", num_blank))

# Calculate the percentage of blank
percentage_blank <- (num_blank / nrow(data)) * 100
print(paste("Percentage of blank:", percentage_blank, "%"))

```
## 1.2 handle with empty strings

```{r}
# Use conditional indexing to get corresponding Type_of_Loan values
non_loan <- data$Type_of_Loan[data$Num_of_Loan <= 0]
head(non_loan)
```

It has been observed that when the values in the Num_of_Loan column are 0 or negative, there still exists some content in the Type_of_Loan column(e.g "Not Specified"). It is considered that a loan count of 0 or negative effectively means there is no loan type. Therefore, the values in the Type_of_Loan column corresponding to 0 or negative values in the Num_of_Loan column should be set to "No_Loan".

```{r}
# Set Type_of_Loan to "Non Loan" where Num_of_Loan is 0 or negative
data$Type_of_Loan <- ifelse(data$Num_of_Loan <= 0, "No Loan", data$Type_of_Loan)
```

## 1.3 distribution of each type

```{r}
# Remove all "and "
data$Type_of_Loan <- gsub("and ", "", data$Type_of_Loan)

# Split by ", "
splitLoans <- strsplit(data$Type_of_Loan, ", ", fixed = TRUE)

# Flatten the list to a vector
loanTypes <- unlist(splitLoans)

# Count the frequency of each text
loanTypeFrequency <- table(loanTypes)

loanTypeFrequencyDF <- as.data.frame(loanTypeFrequency)
colnames(loanTypeFrequencyDF) <- c("LoanType", "Frequency")

library(ggplot2)
ggplot(loanTypeFrequencyDF, aes(x = LoanType, y = Frequency, fill = LoanType)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Frequency), vjust = -0.3, color = "black") +
  theme_minimal() +
  labs(title = "Frequency Distribution of Type_of_Loan", x = "Loan Type", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 25, hjust = 1)) +
  scale_fill_brewer(palette = "Spectral")

```

# 2. Credit_Mix

## 2.1 Count NA and "_" strings

```{r}
data$Credit_Mix <- factor(data$Credit_Mix)
levels_of_Credit_Mix <- levels(data$Credit_Mix)
print(levels_of_Credit_Mix)
```


```{r}
# rename "_" with "Unknown"
#levels(data$Credit_Mix) <- c("Unknown", "Bad", "Good", "Standard")

# Count NA and "_" strings in the Credit_Mix column
num_na <- sum(is.na(data$Credit_Mix))
num_unknown <- sum(data$Credit_Mix == "_")
print(paste("Number of missing values:", num_na))
print(paste("Number of unknown:", num_unknown))

# Calculate the percentage of none
percentage_unknown <- (num_unknown / nrow(data)) * 100
print(paste("Percentage of unknown:", percentage_unknown, "%"))
```

## 2.2 distribution of Credit_Mix

```{r}
frequency_of_Credit_Mix <- table(data$Credit_Mix)
frequencyDF <- as.data.frame(frequency_of_Credit_Mix)
names(frequencyDF) <- c("Level", "Frequency")

ggplot(frequencyDF, aes(x = Level, y = Frequency)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = Frequency), vjust = -0.3, color = "black") + 
  theme_minimal() +
  labs(title = "Distribution of Credit_Mix", x = "Category", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 0, hjust = 1))
```

# 3. Payment_of_Min_Amount

## 3.1 Count NA and NM

```{r}
data$Payment_of_Min_Amount <- factor(data$Payment_of_Min_Amount)
levels_of_Payment_of_Min_Amount <- levels(data$Payment_of_Min_Amount)
print(levels_of_Payment_of_Min_Amount)
```


```{r}
# rename "NM" with "Unknown"
#levels(data$Payment_of_Min_Amount) <- c("Unknown", "No", "Yes")

# Count NA and NM in the Payment_of_Min_Amount column
num_na <- sum(is.na(data$Payment_of_Min_Amount))
num_unknown <- sum(data$Payment_of_Min_Amount == "NM")
print(paste("Number of missing values:", num_na))
print(paste("Number of unknown:", num_unknown))

# Calculate the percentage of none
percentage_unknown <- (num_unknown / nrow(data)) * 100
print(paste("Percentage of unknown:", percentage_unknown, "%"))
```


## 3.3 distribution of Payment_of_Min_Amount

```{r}
frequency_of_Payment_of_Min_Amount <- table(data$Payment_of_Min_Amount)
frequencyDF <- as.data.frame(frequency_of_Payment_of_Min_Amount)
names(frequencyDF) <- c("Level", "Frequency")

ggplot(frequencyDF, aes(x = Level, y = Frequency)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = Frequency), vjust = -0.3, color = "black") + 
  theme_minimal() +
  labs(title = "Distribution of Payment_of_Min_Amount", x = "Category", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 0, hjust = 1))
```


# 4. Payment_Behaviour

## 4.1 Count NA and "!@9#%8"

```{r}
data$Payment_Behaviour <- factor(data$Payment_Behaviour)
levels_of_Payment_Behaviour <- levels(data$Payment_Behaviour)
print(levels_of_Payment_Behaviour)
```

```{r}
# rename "!@9#%8" with "Unknown"
#levels(data$Payment_Behaviour) <- c("Unknown", "High_Large", "High_Medium", "High_Small", "Low_Large", "Low_Medium", "Low_Small")

# Count NA and "!@9#%8" in the Payment_Behaviour column
num_na <- sum(is.na(data$Payment_Behaviour))
num_unknown <- sum(data$Payment_Behaviour == "!@9#%8")
print(paste("Number of missing values:", num_na))
print(paste("Number of unknown:", num_unknown))

# Calculate the percentage of unknown
percentage_unknown <- (num_unknown / nrow(data)) * 100
print(paste("Percentage of unknown:", percentage_unknown, "%"))
```


## 4.2 distribution of Payment_Behaviour

```{r}
frequency_of_Payment_Behaviour <- table(data$Payment_Behaviour)
frequencyDF <- as.data.frame(frequency_of_Payment_Behaviour)
names(frequencyDF) <- c("Level", "Frequency")

ggplot(frequencyDF, aes(x = Level, y = Frequency, fill = Level)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Frequency), vjust = -0.3, color = "black") +
  theme_minimal() +
  labs(title = "Distribution of Payment_Behaviour", x = "Category", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 18, hjust = 1)) +
  scale_fill_viridis_d()
```


# 5. Credit_Score

## 5.1 Count missing value

```{r}
data$Credit_Score <- factor(data$Credit_Score)
levels_of_Credit_Score <- levels(data$Credit_Score)
print(levels_of_Credit_Score)

# Count missing value
num_na <- sum(is.na(data$Payment_Behaviour))
print(paste("Number of missing values:", num_na))

```


## 5.2 distribution of Credit_Score

```{r}
frequency_of_Credit_Score <- table(data$Credit_Score)
frequencyDF <- as.data.frame(frequency_of_Credit_Score)
names(frequencyDF) <- c("Level", "Frequency")

ggplot(frequencyDF, aes(x = Level, y = Frequency, fill = Level)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Frequency), vjust = -0.3, color = "black") +
  theme_minimal() +
  labs(title = "Distribution of Credit_Score", x = "Category", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 18, hjust = 1)) +
  scale_fill_viridis_d()
```


```{r}
count_occupation <- unique(data[data$Occupation != '_______', which(names(data) %in% c("Customer_ID", "Occupation"))] %>% group_by(Customer_ID, Occupation))
sum(duplicated(count_occupation$Customer_ID))
```

```{r}
count_payment_of_min_amount <- unique(data[data$Payment_of_Min_Amount != 'NM', which(names(data) %in% c("Customer_ID", "Payment_of_Min_Amount"))] %>% group_by(Customer_ID, Payment_of_Min_Amount))
sum(duplicated(count_payment_of_min_amount$Customer_ID))
```
```{r}
count_credit_mix <- unique(data[data$Credit_Mix != '_', which(names(data) %in% c("Customer_ID", "Credit_Mix"))] %>% group_by(Customer_ID, Credit_Mix))
sum(duplicated(count_credit_mix$Customer_ID))
```

```{r}
count_payment_behavior <- data[data$Payment_Behaviour != '!@9#%8', which(names(data) %in% c("Customer_ID", "Payment_Behaviour"))] %>% group_by(Customer_ID, Payment_Behaviour) %>% summarise(total_count = n(), .groups = "drop")
count_payment_max_count_by_cust <- count_payment_behavior %>% group_by(Customer_ID) %>% filter(total_count == max(total_count))  %>% do(head(., n = 1))
sum(duplicated(count_payment_max_count_by_cust$Customer_ID))
```

```{r}
invalid_value_df <- data[data$Payment_Behaviour == '!@9#%8' | data$Payment_of_Min_Amount == 'NM' | data$Occupation == '_______' | data$Credit_Mix == '_',]
```
```{r}
fill_invalid_values = function(row){
  if (row[["Payment_Behaviour"]] == '!@9#%8') {
    row["Payment_Behaviour"] <- count_payment_max_count_by_cust[count_payment_max_count_by_cust["Customer_ID"] == row[["Customer_ID"]],]$Payment_Behaviour
  }
  if(row[["Credit_Mix"]] == '_'){
    row["Credit_Mix"] <- count_credit_mix[count_credit_mix["Customer_ID"] == row[["Customer_ID"]],]$Credit_Mix
  }
  if(row[["Occupation"]] == '_______'){
      row["Occupation"] <- count_occupation[count_occupation["Customer_ID"] == row[["Customer_ID"]],]$Occupation
  }
  if(row[["Payment_of_Min_Amount"]] == 'NM'){
    row["Payment_of_Min_Amount"] <- count_payment_of_min_amount[count_payment_of_min_amount["Customer_ID"] == row[["Customer_ID"]],]$Payment_of_Min_Amount
  }
  return(row)
}
for (i in 1:nrow(invalid_value_df)) {
  invalid_value_df[i, ] <- fill_invalid_values(invalid_value_df[i, ])
}
data[rownames(invalid_value_df), ] <- invalid_value_df
```
```{r}
nrow(data[data$Credit_Mix == '_',])
nrow(data[data$Occupation == '_______',])
nrow(data[data$Payment_of_Min_Amount == 'NM',])
nrow(data[data$Payment_Behaviour == '!@9#%8',])
```

```{r}
str(data)
```
```{r}
sum(is.null(data$Month))
data %>% group_by(Month) %>% tally()
```

# 3 Factorize categorical types

```{r}
data <- data %>%
  mutate(
    Credit_Mix_numeric = case_when(
      Credit_Mix == "Bad" ~ 1,
      Credit_Mix == "Standard" ~ 2,
      Credit_Mix == "Good" ~ 3,
      TRUE ~ NA_real_
    ),
    Payment_of_Min_Amount_numeric = case_when(
      Payment_of_Min_Amount == "No" ~ 0,
      Payment_of_Min_Amount == "Yes" ~ 1,
      TRUE ~ NA_real_
    ),
    Payment_Behaviour_numeric = case_when(
      Payment_Behaviour == "High_spent_Large_value_payments" ~ 6,
      Payment_Behaviour == "High_spent_Medium_value_payments" ~ 5,
      Payment_Behaviour == "High_spent_Small_value_payments" ~ 4,
      Payment_Behaviour == "Low_spent_Large_value_payments" ~ 3,
      Payment_Behaviour == "Low_spent_Medium_value_payments" ~ 2,
      Payment_Behaviour == "Low_spent_Small_value_payments" ~ 1,    
      TRUE ~ NA_real_
    ),
    Credit_Score_numeric = case_when(
      Credit_Score == "Poor" ~ 1,
      Credit_Score == "Standard" ~ 2,
      Credit_Score == "Good" ~ 3,
      TRUE ~ NA_real_
    )
  )


head(data)
```

