---
title: "week9"
output: html_document
date: "2024-04-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
data <- read.csv("C:\\Users\\haoyu\\Desktop\\cleaned_data.csv")
```

## Random forest makes predictions by constructing multiple decision trees, which can reduce the risk of overfitting., Provide more stable prediction results. At the same time, it can handle large-scale datasets well.
```{r}
set.seed(5003)
library(randomForest)
data$Credit_Score <- as.factor(data$Credit_Score)
model <- randomForest(Credit_Score ~ ., data = data, ntree = 500, type = "classification")
```


## Setting a threshold based on Gini importance to select features is an effective method to optimize model performance and efficiency, while ensuring that the features used have the greatest impact on the model.
```{r}

importance <- importance(model)

importance_df <- data.frame(Feature = rownames(importance), MeanDecreaseGini = importance[, "MeanDecreaseGini"])

importance_sorted <- importance_df[order(-importance_df$MeanDecreaseGini),]

print(importance_sorted)
```

Considering the number of features and the size of MeanDecreaseGini comprehensively，
we seclect columns which score MeanDecreaseGini over 2200:

Outstanding_Debt	5847.895		
Interest_Rate	4514.114		
Credit_Mix	4257.434		
Credit_History_Age	3776.069		
Delay_from_due_date	3474.969		
Changed_Credit_Limit	3203.479		
Monthly_Balance	2589.577		
Amount_invested_monthly	2581.752		
Credit_Utilization_Ratio	2478.122		
Num_Credit_Card	2354.268	
Total_EMI_per_month	2348.108		
Num_Credit_Inquiries	2281.634		
Monthly_Inhand_Salary	2234.555		
Annual_Income	2224.773		
Num_of_Delayed_Payment	2208.211	


