---
title: "Group Project"
author: "Group 38"
date: "2024-04-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Load libraries
```{r}
library('tidyverse')
library('readr')
library('dplyr')
library('ggplot2')
library('broom')
library('tinytex')
library('mlbench')
library('caret')
```

## Load dataset

```{r}
data <- as.data.frame(read.csv("credit_train.csv", header = TRUE))
```

## Check duplicate rows
```{r}
sum(duplicated(data))
```

## Remove unnecessary columns
```{r}
data <- data[ , -which(names(data) %in% c("ID", "Name", "SSN"))]
```

## Change value for Credit_History_Age column

```{r}
data["Credit_History_Age"] = lapply(data["Credit_History_Age"], function(x) {
    years <- as.integer(sub("\\Years.*", "", x))
    months <- as.integer(sub(".*\\d+\\s+Years\\s+and\\s+(\\d+)\\s+Months.*", "\\1", x))
    return(years*12 + months)
})
```
# Cleaning
## Delete "_"
```{r}
for(i in c("Age","Annual_Income", "Num_of_Loan", "Num_of_Delayed_Payment", 
                    "Changed_Credit_Limit", "Outstanding_Debt", "Amount_invested_monthly", "Monthly_Balance")) {
  data[[i]] <- gsub("_", "", data[[i]])
  
  data[[i]] <- as.numeric(data[[i]])
}

find_most <- function(x) {
  ux <- unique(na.omit(x))
  ux[which.max(tabulate(match(x, ux)))]
}
```
```{r}
colSums(is.na(data))
```
## Age
```{r}
# This is a function to find the bound of the column by using IQR
calculate_bounds <- function(column_name) {
  Q1 <- quantile(data[[column_name]], 0.25, na.rm = TRUE)
  Q3 <- quantile(data[[column_name]], 0.75, na.rm = TRUE)
  IQR <- Q3 - Q1
  lower_bound <- Q1 - 1.5 * IQR
  upper_bound <- Q3 + 1.5 * IQR
  return(c(lower_bound, upper_bound))
}
```
```{r}
# This is a function to check if each group has a unique value
analyze_unique_values <- function(income_column) {
  lower_bound <- calculate_bounds(income_column)[1]
  upper_bound <- calculate_bounds(income_column)[2]
  
  filtered_data <- data %>%
    filter(!!sym(income_column) >= lower_bound & !!sym(income_column) <= upper_bound)
  
  result <- filtered_data %>%
    group_by(Customer_ID) %>%
    summarise(Unique_Value_Count = n_distinct(!!sym(income_column)),
              Only_One_Unique_Value = Unique_Value_Count == 1) %>%
    ungroup()

  if(all(result$Only_One_Unique_Value)) {
    cat("All groups have only one unique value.\n")
  } else {
    cat("These groups have different values:\n")
    print(result %>% filter(Only_One_Unique_Value == FALSE))
  }
}
```

```{r}
analyze_unique_values("Age")
```
```{r}
# This is a function to change outliers with neighbor
clean_data_with_neighbor <- function(column_name, lower_bound, upper_bound) {
  column_sym <- sym(column_name)
  result <- data %>%
    group_by(Customer_ID) %>%
    mutate(
      !!column_sym := ifelse(
        .data[[column_name]] < lower_bound | .data[[column_name]] > upper_bound, NA, .data[[column_name]]
      )
    ) %>%
    fill(!!column_sym, .direction = "up") %>%
    fill(!!column_sym, .direction = "down") %>%
    ungroup()
  
  return(result)
}
```
```{r}
lower_bound <- calculate_bounds("Age")[1]
upper_bound <- calculate_bounds("Age")[2]

data <- clean_data_with_neighbor("Age", 0, upper_bound)
```


## Annual_Income
```{r}
analyze_unique_values("Annual_Income")
```
```{r}
# This is a function that can find the most common value in every column
find_mode <- function(x) {
  ux <- unique(na.omit(x))
  ux[which.max(tabulate(match(x, ux)))]
}
```
```{r}
# This is a function to change outliers with mode
clean_data_with_mode <- function(column_name) {
  col_sym <- rlang::sym(column_name) 
  result <- data %>%
    group_by(Customer_ID) %>%
    mutate(Mode = find_mode(!!col_sym),
           !!col_sym := Mode) %>% 
    select(-Mode) %>%
    ungroup()
  
  return(result)
}
```
```{r}
data <- clean_data_with_mode("Annual_Income")
```
## Monthly_Inhand_Salary

```{r}
# This is a function to fill NA with neighbor
fill_NA_with_neighbour <- function(column_name) {
  result <- data %>%
    group_by(Customer_ID) %>%
    fill(!!rlang::sym(column_name), .direction = "up") %>%
    fill(!!rlang::sym(column_name), .direction = "down") %>%
    ungroup()
  
  return(result)
}
```
```{r}
data <- fill_NA_with_neighbour("Monthly_Inhand_Salary")
```
## Num_Bank_Accounts
```{r}
analyze_unique_values("Num_Bank_Accounts")
```
```{r}
lower_bound <- calculate_bounds("Num_Bank_Accounts")[1]
upper_bound <- calculate_bounds("Num_Bank_Accounts")[2]

data <- clean_data_with_neighbor("Num_Bank_Accounts", 0, upper_bound)
```


## Num_Credit_Card
```{r}
analyze_unique_values("Num_Credit_Card")
```
```{r}
lower_bound <- calculate_bounds("Num_Credit_Card")[1]
upper_bound <- calculate_bounds("Num_Credit_Card")[2]

data <- clean_data_with_neighbor("Num_Credit_Card", 0, upper_bound)
```


## Interest_Rate
```{r}
analyze_unique_values("Interest_Rate")
```
```{r}
data <- clean_data_with_mode("Annual_Income")
```


## Num_of_Loan
```{r}
analyze_unique_values("Num_of_Loan")
```
```{r}
data <- clean_data_with_mode("Num_of_Loan")
```


## Delay_from_due_date
```{r}
# This is a function to replace negative value with 0
replace_negative_with_0 <- function(column_name) {
  result <- data %>%
    mutate(!!rlang::sym(column_name) := ifelse(!!rlang::sym(column_name) < 0, 0, !!rlang::sym(column_name)))
  
  return(result)
}
```
```{r}
data <- replace_negative_with_0("Delay_from_due_date")
```

## Num_of_Delayed_Payment
```{r}
data <- fill_NA_with_neighbour("Num_of_Delayed_Payment")
lower_bound <- calculate_bounds("Num_of_Delayed_Payment")[1]
upper_bound <- calculate_bounds("Num_of_Delayed_Payment")[2]

data <- clean_data_with_neighbor("Num_of_Delayed_Payment", 0, upper_bound)
```


## Changed_Credit_Limit
```{r}
data <- replace_negative_with_0("Changed_Credit_Limit")
```
```{r}
# This is a function to fill NA with mean
fill_NA_with_mean <- function(column_name) {
  col_sym <- rlang::sym(column_name)
  result <- data %>%
    group_by(Customer_ID) %>%
    mutate(
      Mean = round(mean({{col_sym}}[!is.na({{col_sym}})], na.rm = TRUE), 2),
      {{col_sym}} := if_else(is.na({{col_sym}}), Mean, {{col_sym}})
    ) %>%
    select(-Mean) %>%
    ungroup()
  return(result)
}
```
```{r}
data <- fill_NA_with_mean("Changed_Credit_Limit")
```


## Num_Credit_Inquiries
```{r}
analyze_unique_values("Num_Credit_Inquiries")
```
```{r}
data <- fill_NA_with_neighbour("Num_Credit_Inquiries")
lower_bound <- calculate_bounds("Num_Credit_Inquiries")[1]
upper_bound <- calculate_bounds("Num_Credit_Inquiries")[2]

data <- clean_data_with_neighbor("Num_Credit_Inquiries", 0, upper_bound)
```


## Outstanding_Debt
```{r}
analyze_unique_values("Outstanding_Debt")
```
```{r}
data <- clean_data_with_mode("Outstanding_Debt")
```


## Credit_Utilization_Ratio
```{r}
out_of_range <- data %>%
  filter(Credit_Utilization_Ratio < 0 | Credit_Utilization_Ratio > 100)
print(out_of_range)
if(nrow(out_of_range) > 0) {
  print("There are values outside the 0 to 100 range")
} else {
  print("All values are in 0 to 100 range")
}
```


## Credit_History_Age
```{r}
data <- data %>%
  group_by(Customer_ID) %>%
  mutate(Index = row_number()) %>%
  arrange(Customer_ID, Index) %>%
  mutate(First_Valid = min(Index[!is.na(Credit_History_Age)])) %>% 
  mutate(Base = Credit_History_Age[First_Valid] - First_Valid) %>% 
  mutate(Credit_History_Age = if_else(is.na(Credit_History_Age), Index + Base, Credit_History_Age)) %>%
  select(-Index, -First_Valid, -Base) %>%
  ungroup()
```


## Total_EMI_per_month

```{r}
analyze_unique_values("Total_EMI_per_month")
```
```{r}
lower_bound <- calculate_bounds("Total_EMI_per_month")[1]
upper_bound <- calculate_bounds("Total_EMI_per_month")[2]

data <- clean_data_with_neighbor("Total_EMI_per_month", 0, 1800)
```


## Amount_invested_monthly
```{r}
unique(data$Amount_invested_monthly)
```
```{r}
# This is a function to replace the outliers and NA with mean
replace_with_group_mean <- function(column_name, value_to_replace) {
  col_sym <- sym(column_name)
  result <- data %>%
    group_by(Customer_ID) %>%
    mutate(
      Group_Mean = mean({{col_sym}}[{{col_sym}} != value_to_replace & !is.na({{col_sym}})], na.rm = TRUE),
      {{col_sym}} := if_else({{col_sym}} == value_to_replace | is.na({{col_sym}}), Group_Mean, {{col_sym}})
    ) %>%
    select(-Group_Mean) %>%
    ungroup()
  return(result)
}
```
```{r}
data <- replace_with_group_mean("Amount_invested_monthly", 10000)
```


## Monthly_Balance
```{r}
unique(data$Monthly_Balance)
```
```{r}
data <- data %>%
  mutate(Monthly_Balance = ifelse(Monthly_Balance == -333333333333333333333333333, 333333333333333333333333333, Monthly_Balance))
```
```{r}
data <- replace_with_group_mean("Monthly_Balance", 333333333333333333333333333)
```
```{r}
colSums(is.na(data))
```
```{r}
data
```

