---
title: "Group Project"
author: "Group 38"
date: "2024-04-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Load libraries
```{r}
library('tidyverse')
library('readr')
library('dplyr')
library('ggplot2')
library('broom')
library('tinytex')
library('mlbench')
library('caret')
```

## Load dataset

```{r}
data <- as.data.frame(read.csv("credit_train.csv", header = TRUE))
```

## Check duplicate rows
```{r}
sum(duplicated(data))
```

## Remove unnecessary columns
```{r}
data <- data[ , -which(names(data) %in% c("ID", "Name", "SSN"))]
```

## Change value for Credit_History_Age column

```{r}
data["Credit_History_Age"] = lapply(data["Credit_History_Age"], function(x) {
    years <- as.integer(sub("\\Years.*", "", x))
    months <- as.integer(sub(".*\\d+\\s+Years\\s+and\\s+(\\d+)\\s+Months.*", "\\1", x))
    return(years*12 + months)
})
```
```{r}
str(data)
```
```{r}
data
```



# Missing values {.tabset .tabset-fade .tabset-pills}
## Count missing values {.tabset .tabset-fade .tabset-pills}

```{r}
colSums(is.na(data))
```

```{r}
empty_string_counts <- data %>%
  summarise(across(everything(), ~ sum(.x == "", na.rm = TRUE))) %>%
  select_if(~ .x > 0)
print(empty_string_counts)
```


## Monthly_Inhand_Salary {.tabset .tabset-fade .tabset-pills}

```{r}
data <- data %>%
  group_by(Customer_ID) %>%
  fill(Monthly_Inhand_Salary, .direction = "up") %>%
  fill(Monthly_Inhand_Salary, .direction = "down") %>%
  ungroup()
```

```{r}
# Test
data
```
## Num_Credit_Inquiries {.tabset .tabset-fade .tabset-pills}
```{r}
data <- data %>%
  group_by(Customer_ID) %>%
  fill(Num_Credit_Inquiries, .direction = "up") %>%
  fill(Num_Credit_Inquiries, .direction = "down") %>%
  ungroup()
```

```{r}
data <- data %>%
  group_by(Customer_ID) %>%
  mutate(Q1 = quantile(Num_Credit_Inquiries, 0.25, na.rm = TRUE),
         Q3 = quantile(Num_Credit_Inquiries, 0.75, na.rm = TRUE),
         IQR = Q3 - Q1,
         Lower_Bound = Q1 - 1.5 * IQR,
         Upper_Bound = Q3 + 1.5 * IQR) %>%
  mutate(Num_Credit_Inquiries = if_else(Num_Credit_Inquiries < Lower_Bound | 
                                        Num_Credit_Inquiries > Upper_Bound, NA_real_, Num_Credit_Inquiries)) %>%
  select(-Q1, -Q3, -IQR, -Lower_Bound, -Upper_Bound) %>%
  fill(Num_Credit_Inquiries, .direction = "up") %>%
  fill(Num_Credit_Inquiries, .direction = "down") %>%
  ungroup()
```

```{r}
# Test
data
```


## Credit_History_Age {.tabset .tabset-fade .tabset-pills}
```{r}
data <- data %>%
  group_by(Customer_ID) %>%
  mutate(Index = row_number()) %>%
  arrange(Customer_ID, Index) %>%
  mutate(First_Valid = min(Index[!is.na(Credit_History_Age)])) %>% 
  mutate(Base = Credit_History_Age[First_Valid] - First_Valid) %>% 
  mutate(Credit_History_Age = if_else(is.na(Credit_History_Age), Index + Base, Credit_History_Age)) %>%
  select(-Index, -First_Valid, -Base) %>%
  ungroup()
```

```{r}
# Test
colSums(is.na(data))
```

```{r}
filtered_data <- data %>%
  filter(Customer_ID == "CUS_0x3e2d")
print(filtered_data)
```


## Num_Bank_Accounts {.tabset .tabset-fade .tabset-pills}
```{r}
# This is a function that can find the most common value in every column
find_most <- function(x) {
  ux <- unique(na.omit(x))
  ux[which.max(tabulate(match(x, ux)))]
}
```
```{r}
result <- data %>%
  group_by(Customer_ID) %>%
  mutate(Mode = find_most(Num_Bank_Accounts),
         Num_Bank_Accounts = Mode) %>% 
  select(-Mode) %>% 
  ungroup() 
```


## Num_Credit_Card {.tabset .tabset-fade .tabset-pills}
```{r}
result <- data %>%
  group_by(Customer_ID) %>%
  mutate(Mode = find_most(Num_Credit_Card),
         Num_Credit_Card = Mode) %>% 
  select(-Mode) %>% 
  ungroup() 
```


## Interest_Rate {.tabset .tabset-fade .tabset-pills}
```{r}
result <- data %>%
  group_by(Customer_ID) %>%
  mutate(Mode = find_most(Interest_Rate),
         Interest_Rate = Mode) %>% 
  select(-Mode) %>% 
  ungroup() 
```


## Credit_Utilization_Ratio {.tabset .tabset-fade .tabset-pills}
```{r}
out_of_range <- data %>%
  filter(Credit_Utilization_Ratio < 0 | Credit_Utilization_Ratio > 100)
print(out_of_range)
if(nrow(out_of_range) > 0) {
  print("There are values outside the 0 to 100 range")
} else {
  print("All values are in 0 to 100 range")
}
```


## Total_EMI_per_month {.tabset .tabset-fade .tabset-pills}
```{r}
data <- data %>%
  group_by(Customer_ID) %>%
  mutate(Q1 = quantile(Total_EMI_per_month, 0.25, na.rm = TRUE),
         Q3 = quantile(Total_EMI_per_month, 0.75, na.rm = TRUE),
         IQR = Q3 - Q1,
         Lower_Bound = Q1 - 1.5 * IQR,
         Upper_Bound = Q3 + 1.5 * IQR) %>%
  mutate(Total_EMI_per_month = if_else(Total_EMI_per_month < Lower_Bound | 
                                        Total_EMI_per_month > Upper_Bound, NA_real_, Total_EMI_per_month)) %>%
  select(-Q1, -Q3, -IQR, -Lower_Bound, -Upper_Bound) %>%
  fill(Total_EMI_per_month, .direction = "up") %>%
  fill(Total_EMI_per_month, .direction = "down") %>%
  ungroup()
```

```{r}
data
```




